---
title: "The Use of Force by Police"
author: "Anthony Lee, Niranjanaa Kannan, Munir Emam, Dev Dhawan"
date: "11/14/2022"
output: html_document
---

```{r setup, include=FALSE}
source("../source/Chart1.R")
source("../source/Table.R")
source("../source/Chart2.R")
source("../source/Chart3.R")
knitr::opts_chunk$set(echo = TRUE)
```

## Dynamic Paragraph

The data sets pertain to use of force by police officers, aggregated by geographic regions, gender, and race. In our findings, the mode for gender subjected to use of force was seen to be males, with 11007 cases of male victims in Seattle compared to 3139 female victims in Seattle. In addition there was 2779 male victims in Dallas, compared to only 590 female victims in Dallas. This is a stark skew in the data set, as the most common gender in use of force cases pertains to males by several times more than female.

The data sets also records the race of victims in incidents involving the use of force. The races recorded in the Seattle data set were Black, White, Asian, Native American, Native Hawaiian / Pacific Islander, Not Specified, Native American, and Hispanic or Latino. The races recorded in Dallas were Black, White, Asian, Hispanic, Other, NULL, and Unknown. The race counts let us analyze whether certain groups are more prone to be victims of the use of force.

The New Orleans data set provided information on outside factors involved in the use of force incidents. Outside factors recorded include mentally unstable, alcohol, unknown, unknown drugs, marijuana, incoherent, agitated, animal, alcohol and unknown drugs, talking to themselves, poor hygiene, NULL, and none detected. These factors allowed us to further analyze whether certain groups of people are more likely to be involved in use of force incidents. 

To try and discover how many incidents are deemed as unjustified, the New Orleans data set provides information on the number of incidents that the New Orleans Police Department deems as unjustified. Based on the data set, the number of unjustified incidents was 48 equating to a proportion of 0.016. However, there is limited information on what is unjustified and it is based off the discretion of the police departments investigation.

These visual summaries of the overall use of force data allow one to put forth data-driven analysis on specific factors within use-of-force incidents and make data-driven conclusions. However, there are still some limitations, including context for overall population and further segmentation of factors that impact use of force. Nonetheless, the data still points towards numbers that can be used to reasonably identify key factors in use of force incidents.

## Table
```{r echo=FALSE}
knitr::kable(df, "pipe", col.name = c("Race", "Seattle Race Count", "Race", "Dallas Race Count", "Race", "New Orleans Race Count"), )
```
The table shows the information about subjects who experienced the use of force by law enforcement in the city of Seattle, Dallas, and New Orleans. The table was included to show the distribution of subjects who experienced the use of force by law enforcement based on race. Specifically, it was used to examine whether specific groups are more likely to be subjects of the use of force. The table shows that whites account for a large majority of subjects with blacks being the second largest group. Groups like native american make up a small proportion of subjects.

## Graph 1
```{r echo=FALSE}
chart
```

The purpose of the pie chart is to show the percentage of force being used on a subject based on race. The pie chart allows us to identify trends in which groups are more likely to be subject of the use of force by law enforcement. Observations that can be made from the pie chart is that Blacks or African Americans and Whites account for a large majority of subjects in incidents involving the use of force. However, the Blacks or African Americans account for a significantly less proportion of the US population when compared to Whites. This may point to systematic racism.

## Graph 2
``` {r echo=FALSE}
barplot
```

This bar chart shows the total use of force with the officer's years in service. This bar chart lets us examine the trends between which officers will have a higher likelyhood in using force, and which ones don't. After observing this bar chart it shows that officers with 7-9 years of service been involved in the most use of force incidents. This bar chart also shows that the more years of service, the less use of force incidents there are, which may point that more training/experience leads to less use of force. 


## Graph 3
``` {r echo=FALSE}
factors_graph
```

The pack circle chart shows the number of outside factors involved in the use of force. We included the chart because it lets us see what the most common factors contribute to the likelyhood of force being used. The size of the circles based on the number of recorded incidents allows for viewers to easily determine the largest contributing factors. The packed circle chart shows us that mental illness, alcohol, and drugs are key factors that significantly raise the likelyhood of force being involved. 