---
title: "The Use of Force by Police"
author: "Anthony Lee, Niranjanaa Kannan, Munir Emam, Dev Dhawan"
date: "11/14/2022"
output: html_document
---

```{r setup, include=FALSE}
source("../source/Chart1.R")
source("../source/Table.R")
source("../source/Chart2.R")
source("../source/Chart3.R")
knitr::opts_chunk$set(echo = TRUE)
```

## Dynamic Paragraph

The data sets pertain to use of force by police officers, aggregated by geographic region and gender, then by race. The following charts include a table and pie chart to provide various visualizations of the data, each focusing on a different aspect of use of force incidents.

The table aggregates the data by separating the victims of use of force by gender, to develop and data-based correlation between gender and use of force by police. In our findings, the mode for gender subjected to use of force was seen to be males, with 11007 cases of male victims compared to 3114 female victims, and 25 transgender female and other genders to follow. This is a stark skew in the data set, as the most common gender in use of force cases pertains to males by several times more than female, transgender females and males, and so on forth.

The pie chart provides a visual summary of the victims of use of force by race, to bring forth a conclusion on the correlation between race and police use of force. In our findings the max percent of use of force incidents were subjected to White victims (making up 40% of the chart). However, 58.3% of the U.S. population is white. In the pie chart, 33% of use of force incidents were subjected to Black victims. However, only 12.6% of the U.S. population is Black. Based on this information, it can be reasonably concluded that Black people are disproportionately impacted by use-of force incidents, and are far more likely to encounter this issue than the average white resident of America.

These visual summaries of the overall use of force data allow one to put forth data-driven analysis on specific factors within use-of-force incidents and make data-driven conclusions. However, there are still some limitations, including context for overall population and further segmentation of factors that impact use of force. Nonetheless, the data still points towards numbers that can be used to reasonably identify key factors in use of force incidents.

## Table
```{r echo=FALSE}
knitr::kable(df, "pipe", col.name = c("Race", "Seattle Race Count", "Race", "Dallas Race Count", "Race", "New Orleans Race Count"), )
```
The table shows the information about subjects who experienced the use of force by law enforcement in the city of Seattle, Dallas, and New Orleans. The table was included to show the distribution of subjects who experienced the use of force by law enforcement based on race. Specifically, it was used to examine whether specific groups are more likely to be subjects of the use of force. The table shows that whites account for a large majority of subjects with blacks being the second largest group. Groups like native american make up a small proportion of subjects.

## Graph 1
```{r echo=FALSE}
chart
```

The purpose of the pie chart is to show the percentage of force being used on a subject based on race. The pie chart allows us to identify trends in which groups are more likely to be subject of the use of force by law enforcement. Observations that can be made from the pie chart is that Blacks or African Americans and Whites account for a large majority of subjects in incidents involving the use of force. However, the Blacks or African Americans account for a significantly less proportion of the US population when compared to Whites. This may point to systematic racism.

## Graph 2
``` {r echo=FALSE}
barplot
```

This bar chart shows the total use of force with the officer's years in service. This bar chart lets us examine the trends between which officers will have a higher likelyhood in using force, and which ones don't. After observing this bar chart it shows that officers with 7-9 years of service been involved in the most use of force incidents. This bar chart also shows that the more years of service, the less use of force incidents there are, which may point that more training/experience leads to less use of force. 


## Graph 3
``` {r echo=FALSE}
factors_graph
```

The pack circle chart shows the number of outside factors involved in the use of force. We included the chart because it lets us see what the most common factors contribute to the likelyhood of force being used. The size of the circles based on the number of recorded incidents allows for viewers to easily determine the largest contributing factors. The packed circle chart shows us that mental illness, alcohol, and drugs are key factors that significantly raise the likelyhood of force being involved. 